### 2016-05-16 text/AppendixB.tex
Spell checking

This appendinx provides additional details for some of the proofs for equational theory
of flat coeffect calculus from Chapter~\ref{ch:flat} and structural coeffect calculus
from Chapter~\ref{ch:structural}.

>>>>

This appendix provides additional details for some of the proofs for equational theory
of flat coeffect calculus from Chapter~\ref{ch:flat} and structural coeffect calculus
from Chapter~\ref{ch:structural}.

### 2016-05-16 text/Chapter02.tex
Spell checking

This chapter serves two purposes. Firstly, it provides a high-level overview of the  related work,
although technical details are often postponed until later. Secondly it recasts existing ideas in
a way that naturally leads to the coeffect systems developed later in the thesis. For this reason,
we are not always faithful to the referenced work -- sometimes we focus on aspects that the
authors consider unimportant or present the work differently than originally intended. The reason
is to fulfil the second goal of the chapter. When we do so, this is explicitly said in the text.

@@

In static and dynamic scoping, all variable requirements are always placed on one site. However,
those are not the only two options. It is conceivable that a language would use a mechanism that
splits variable requirements differently and combines aspects of dynamic and static binding. For
example, the language could use statis binding by default, but resort to dynamic binding if a
variable is not available in the lexical scope. One such system is discussed in the next section.

@@

\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\kvd{let}~\ident{?x}=10~\kvd{in}\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{?x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{?x}=5~\kvd{in}\\
\ident{f}~0
\end{array}\hspace{5em}\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{?x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{?x}=5~\kvd{in}\\
\ident{f}~0
\end{array}
\end{equation*}
%
The binding rules for implicit parameters can be summarized as ``\emph{static binding when possible,
dynamic binding when needed}''. If an implicit parameter is available in scope, then the value
is statically bound and the context requirement is satisfied using the declaration-site context.
Otherwise, the context requirement is delayed and has to be satisfied at the call site.
In the example on the left, \ident{?x} is bound to $10$ the function \ident{f} has no
delayed context requirements and the expression evalautes to $10$. In the example on the right,
the context requirement \ident{?x} is delayed and is satisfied via dynamic binding when calling
the function. The expression evaluates to $5$.

>>>>

This chapter serves two purposes. Firstly, it provides a high-level overview of the  related work,
although technical details are often postponed until later. Secondly it recasts existing ideas in
a way that naturally leads to the coeffect systems developed later in the thesis. For this reason,
we are not always faithful to the referenced work -- sometimes we focus on aspects that the
authors consider unimportant or present the work differently than originally intended. The reason
is to fulfill the second goal of the chapter. When we do so, this is explicitly said in the text.

@@

In static and dynamic scoping, all variable requirements are always placed on one site. However,
those are not the only two options. It is conceivable that a language would use a mechanism that
splits variable requirements differently and combines aspects of dynamic and static binding. For
example, the language could use static binding by default, but resort to dynamic binding if a
variable is not available in the lexical scope. One such system is discussed in the next section.

@@

\begin{equation*}
\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\kvd{let}~\ident{?x}=10~\kvd{in}\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{?x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{?x}=5~\kvd{in}\\
\ident{f}~0
\end{array}\hspace{5em}\begin{array}{l}
\kvd{let}~\ident{f}=\\[-0.25em]
\qquad\lambda \ident{y}\rightarrow \ident{?x}+\ident{y}~\kvd{in}\\[-0.25em]
\kvd{let}~\ident{?x}=5~\kvd{in}\\
\ident{f}~0
\end{array}
\end{equation*}
%
The binding rules for implicit parameters can be summarized as ``\emph{static binding when possible,
dynamic binding when needed}''. If an implicit parameter is available in scope, then the value
is statically bound and the context requirement is satisfied using the declaration-site context.
Otherwise, the context requirement is delayed and has to be satisfied at the call site.
In the example on the left, \ident{?x} is bound to $10$ the function \ident{f} has no
delayed context requirements and the expression evaluates to $10$. In the example on the right,
the context requirement \ident{?x} is delayed and is satisfied via dynamic binding when calling
the function. The expression evaluates to $5$.

### 2016-05-16 text/Chapter03.tex
Spell checking

When introducing coeffect systems in Section~\ref{sec:path-effects-coeff}, we related coeffect systems
with effect systems. Effect systems track how a program affects the environment, or, in other words
capture some \emph{output impurity}. In contrast, coeffect systems track what a program requires from
the execution envionment, or \emph{input impurity}.

@@

In (\emph{var}), the context contains just a single variable and so we do not even need to apply
projection; (\emph{cosnt}) receives no variables and uses global constant lookup function $\delta$.
In (\emph{abs}), we obtain two parts of the context and combine them into $(\mathbf{v}, x)$. This
works the same way regardless of whether the variables are live or dead. For simplicity, we omit
subcoeffecting, which just turns some of the available values $v_i$ to unit values $()$.

@@

In the following, we use vector names $\mathbf{v}_i$ for contexts containing multiple variables
\ie~have a type $\tau_1^{\cclrd{r_1}}\times\ldots\times\tau_m^{\cclrd{r_m}}$. Each vector contains
multiple copies of each variable, to model the fact that variables are used in an affine way (at most
once). We do not explicitly write the sizes of these vectors (number of variables in a context; number
of instances of a variable) as these are clear from the coeffect annotations. We assume that $\Gamma_2$
contains $n$ variables and that $\aclrd{s}=\alift{\cclrd{s}_1, \ldots, \cclrd{s}_n}$. First, consider
the (\emph{contr}) rule:
%
\begin{equation*}
\hspace{-1em}
\begin{array}{l}
\mdeff
  {\begin{array}{l}\sem{\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2\\[-0.25em]\qquad}
      {\aclrd{\textbf{r}} \atimes \alift{\cclrd{s},\cclrd{t}} \atimes \aclrd{\textbf{q}}} \vdash e : \tau}\end{array}}
  {\begin{array}{l}\sem{\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}}
          \atimes \alift{\cclrd{s} + \cclrd{t}} \atimes \aclrd{\textbf{q}}} \\[-0.25em] \qquad\vdash \subst{e}{z,y}{x} : \tau}\end{array}}
  {\begin{array}{l}~\\[-0.25em]\hspace{-0.5em}f\end{array}}
  {\hspace{-0.5em}\begin{array}{l}\lambda(\mathbf{v_1}, (x_1, \ldots, x_{\cclrd{s}+\cclrd{t}}), \mathbf{v_2}) .\\[-0.25em]
  \quad f~(\mathbf{v_1}, (x_1, \ldots, x_{\cclrd{s}}), (x_{\cclrd{s}+1}, \ldots, x_{\cclrd{s}+\cclrd{t}} ), \mathbf{v_2})\end{array}\hspace{-0.5em}~}
\end{array}
\end{equation*}
%
The semantic function is called with $\cclrd{s}+\cclrd{t}$ copies of
a value for the $x$ variable. The values are split between $\cclrd{s}$ and $\cclrd{t}$ separate
copies of variables $y$ and $z$, respectively. The (\emph{app}) rule is similar in that it needs to
split the input variable context. However, it needs to split values of multiple variabless:
%
\begin{equation*}
\hspace{-1em}
\begin{array}{l}
\emdefff
  {\coctx{\Gamma_1}{\aclrd{\textbf{r}}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2}
  {\coctx{\Gamma_2}{\aclrd{\textbf{s}}} \vdash e_2 : \tau_1}
  {\begin{array}{l}\sem{\coctx{\Gamma_1, \Gamma_2}{\aclrd{\textbf{r}} \cons (\cclrd{t} \ast \aclrd{\textbf{s}})}\\[-0.25em]
    \qquad\vdash e_1 \, e_2 : \tau_2}\\[-0.25em]~\\[-0.25em]~\\[-0.25em]~\\[-0.25em]~\end{array}\hspace{-0.5em}~}
  {f_1}
  {f_2}
  {\hspace{-0.5em}\begin{array}{l}
  \lambda (\mathbf{v_1}, ((x_{1,1}, \ldots, x_{1,\cclrd{t}\ast\cclrd{s}_1}), \ldots, (x_{n,1}, \ldots, x_{n,\cclrd{t}\ast\cclrd{s}_n}) ) .\\[-0.25em]
  \quad(f_1~\textbf{v}_1)\\[-0.25em]
  \qquad (~f_2~((x_{1,1}, \ldots, x_{1,\cclrd{s}_1}), \ldots, \\[-0.25em]
  \hspace{4.0em}                                         ~(x_{n,1}, \ldots, x_{1,\cclrd{s}_n}) ~),\ldots,\\[-0.25em]
  \qquad \;~f_2~((x_{1,(\cclrd{t}-1)\ast\cclrd{s}_1 + 1}, \ldots, x_{1,\cclrd{t}\ast\cclrd{s}_1}), ~\ldots~, \\[-0.25em]
  \hspace{4.0em}                                         ~(x_{n,(\cclrd{t}-1)\ast\cclrd{s}_n + 1}, \ldots, x_{1,\cclrd{t}\ast\cclrd{s}_n})  ~)~)\\[-0.25em]
  \end{array}}}
\end{array}
\end{equation*}
\vspace{0.5em}

>>>>

When introducing coeffect systems in Section~\ref{sec:path-effects-coeff}, we related coeffect systems
with effect systems. Effect systems track how a program affects the environment, or, in other words
capture some \emph{output impurity}. In contrast, coeffect systems track what a program requires from
the execution environment, or \emph{input impurity}.

@@

In (\emph{var}), the context contains just a single variable and so we do not even need to apply
projection; (\emph{const}) receives no variables and uses global constant lookup function $\delta$.
In (\emph{abs}), we obtain two parts of the context and combine them into $(\mathbf{v}, x)$. This
works the same way regardless of whether the variables are live or dead. For simplicity, we omit
subcoeffecting, which just turns some of the available values $v_i$ to unit values $()$.

@@

In the following, we use vector names $\mathbf{v}_i$ for contexts containing multiple variables
\ie~have a type $\tau_1^{\cclrd{r_1}}\times\ldots\times\tau_m^{\cclrd{r_m}}$. Each vector contains
multiple copies of each variable, to model the fact that variables are used in an affine way (at most
once). We do not explicitly write the sizes of these vectors (number of variables in a context; number
of instances of a variable) as these are clear from the coeffect annotations. We assume that $\Gamma_2$
contains $n$ variables and that $\aclrd{s}=\alift{\cclrd{s}_1, \ldots, \cclrd{s}_n}$. First, consider
the (\emph{contr}) rule:
%
\begin{equation*}
\hspace{-1em}
\begin{array}{l}
\mdeff
  {\begin{array}{l}\sem{\coctx{\Gamma_1,y\!:\!\tau_1,z\!:\!\tau_1,\Gamma_2\\[-0.25em]\qquad}
      {\aclrd{\textbf{r}} \atimes \alift{\cclrd{s},\cclrd{t}} \atimes \aclrd{\textbf{q}}} \vdash e : \tau}\end{array}}
  {\begin{array}{l}\sem{\coctx{\Gamma_1,x\!:\!\tau_1,\Gamma_2}{\aclrd{\textbf{r}}
          \atimes \alift{\cclrd{s} + \cclrd{t}} \atimes \aclrd{\textbf{q}}} \\[-0.25em] \qquad\vdash \subst{e}{z,y}{x} : \tau}\end{array}}
  {\begin{array}{l}~\\[-0.25em]\hspace{-0.5em}f\end{array}}
  {\hspace{-0.5em}\begin{array}{l}\lambda(\mathbf{v_1}, (x_1, \ldots, x_{\cclrd{s}+\cclrd{t}}), \mathbf{v_2}) .\\[-0.25em]
  \quad f~(\mathbf{v_1}, (x_1, \ldots, x_{\cclrd{s}}), (x_{\cclrd{s}+1}, \ldots, x_{\cclrd{s}+\cclrd{t}} ), \mathbf{v_2})\end{array}\hspace{-0.5em}~}
\end{array}
\end{equation*}
%
The semantic function is called with $\cclrd{s}+\cclrd{t}$ copies of
a value for the $x$ variable. The values are split between $\cclrd{s}$ and $\cclrd{t}$ separate
copies of variables $y$ and $z$, respectively. The (\emph{app}) rule is similar in that it needs to
split the input variable context. However, it needs to split values of multiple variables:
%
\begin{equation*}
\hspace{-1em}
\begin{array}{l}
\emdefff
  {\coctx{\Gamma_1}{\aclrd{\textbf{r}}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2}
  {\coctx{\Gamma_2}{\aclrd{\textbf{s}}} \vdash e_2 : \tau_1}
  {\begin{array}{l}\sem{\coctx{\Gamma_1, \Gamma_2}{\aclrd{\textbf{r}} \cons (\cclrd{t} \ast \aclrd{\textbf{s}})}\\[-0.25em]
    \qquad\vdash e_1 \, e_2 : \tau_2}\\[-0.25em]~\\[-0.25em]~\\[-0.25em]~\\[-0.25em]~\end{array}\hspace{-0.5em}~}
  {f_1}
  {f_2}
  {\hspace{-0.5em}\begin{array}{l}
  \lambda (\mathbf{v_1}, ((x_{1,1}, \ldots, x_{1,\cclrd{t}\ast\cclrd{s}_1}), \ldots, (x_{n,1}, \ldots, x_{n,\cclrd{t}\ast\cclrd{s}_n}) ) .\\[-0.25em]
  \quad(f_1~\textbf{v}_1)\\[-0.25em]
  \qquad (~f_2~((x_{1,1}, \ldots, x_{1,\cclrd{s}_1}), \ldots, \\[-0.25em]
  \hspace{4.0em}                                         ~(x_{n,1}, \ldots, x_{1,\cclrd{s}_n}) ~),\ldots,\\[-0.25em]
  \qquad \;~f_2~((x_{1,(\cclrd{t}-1)\ast\cclrd{s}_1 + 1}, \ldots, x_{1,\cclrd{t}\ast\cclrd{s}_1}), ~\ldots~, \\[-0.25em]
  \hspace{4.0em}                                         ~(x_{n,(\cclrd{t}-1)\ast\cclrd{s}_n + 1}, \ldots, x_{1,\cclrd{t}\ast\cclrd{s}_n})  ~)~)\\[-0.25em]
  \end{array}}}
\end{array}
\end{equation*}
\vspace{0.5em}

### 2016-05-16 text/Chapter04.tex
Spell checking

We enrich types and typing judgements with coeffect annotations $\cclrd{r}, \cclrd{s}, \cclrd{t}$;
typing judgements are written as $\coctx{\Gamma}{\cclrd{r}}\vdash e:\tau$. The expressions of
the calculus are those of the $\lambda$-calculus with \emph{let} binding. We also include a type
\ident{num} as an example of a concrete base type with numerical constants written as $n$:
%
\begin{equation*}
\begin{array}{rcl}
e &::=& x \sep n \sep \lambda x:\tau.e \sep e_1~e_2 \sep \kvd{let}~x = e_1~\kvd{in}~e_2\\
\tau &::=& \ident{num} \sep \tau_1 \xrightarrow{\cclrd{r}} \tau_2
\end{array}
\end{equation*}
%
Note that the lambda abstrction in the syntax is written in the Church-style and requires a type
annotation. This will be used in Section~\ref{sec:flat-unique} where we discuss how to find a
unique typing derivation for context-aware computations. Using Church-style lambda abstraction,
we can directly focus on the more interesting problem of finding unique \emph{coeffect annotations}
rather than solving the problem of type reconstruction.

@@

To guide understanding of the system, we also show non-syntax-directed (\emph{sub}) rule for
subcoeffecting. The rule states that an expression with context demands $\cclrd{r'}$ can be
treated as an expression with greater context demands $\cclrd{r}$. We return to subcoeffecting,
subtyping and additional constructs such as pairs in Section~\ref{sec:flat-exts}. When
discussing procedure for choocing unique typing in Section~\ref{sec:flat-unique}, we consider
only the syntax-directed part of the system.

@@

\paragraph{Properties.}
If a program written in a coeffect language with implicit parameters is well-typed in a type
system presented in Figure~\ref{fig:flat-resolve-impl} then this identifies the unique prefered
derivation for the program. We use this unique typing derivation to give the semantics of coeffect
language with implicit parameters in Chapter~\ref{ch:semantics} and we also implement this algorithm
as discussed in Chatper~\ref{ch:impl}.

@@

\paragraph{Properties.}
If a program written in a coeffect language for liveness or dataflow is well-typed according to
the type system presented in Figure~\ref{fig:flat-types} with the (\emph{abs}) rule replaced by
(\emph{idabs}), then the type system gives a unique derivation. As for implicit parameters, this
defines the semantics of coeffect program (Chapter~\ref{ch:semantics}) and it is used in the
implementation (Chatper~\ref{ch:impl}).

@@

\paragraph{Implicit parameters.} In dataflow, substituting $\kvd{prev}~x$ for a variable $y$ in
an expression $\lambda z.y$ changes the context demands attached to the type of the function.
This is the case not just for the preferred unique typeing derivation, but for all possible typings
that can be obtained using the (\emph{abs}) rule. However, this is not the case for all
systems. Consider a substitution $\subst{\lambda x.y}{y}{\ident{?p}}$ that substitutes
an implicit parameter access $\ident{?p}$ for a free variable $y$ under a lambda:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\emptyset} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{ \{\ident{?p}\} } &\narrow{\vdash}& \lambda x.\ident{?p} : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
The (\emph{after}) judgement shows one possible typing of the body -- one that does not change the
coeffects of the function type and attaches all additional coeffects (implicit parameters) to the
context. In case of implicit parameters (and, more generally, systems with set-like annotations)
this is always possible.

@@

It is perhaps somewhat unexpected that the annotation can be simplified in different ways for
different examples. However, for all our systems, the simplfied annotation (right column in
Table~\ref{tab:flat-simplelet}) is more precise than the original (left column). Recall that
$\cclrd{s} = \cclrd{s_1}\,\czip\,\cclrd{s_2}$. The following inequalities hold:
%
\begin{equation*}
\begin{array}{rclll}
\cclrd{s_1} \cup (\cclrd{s_2} \cup \cclrd{r}) &\supseteq& (\cclrd{s_1} \cup \cclrd{s_2}) \cup \cclrd{r}
  && \textnormal{(implicit parameters)}\\
\cclrd{s_1} \sqcap (\cclrd{s_2} \sqcup \cclrd{r}) &\sqsupseteq&  (\cclrd{s_1} \sqcap \cclrd{s_2})
  && \textnormal{(liveness)} \\
\mathit{max}(\cclrd{s_1}, \cclrd{s_2} + \cclrd{r}) &\geq& \mathit{min}(\cclrd{s_1}, \cclrd{s_2}) + \cclrd{r}
  &\quad& \textnormal{(dataflow)} \\
\end{array}
\end{equation*}
%
In other words, the inequality states that using idempotence, we get a more precise typing.
Using the $\cgeq$ operator of flat coeffect algebra, this property can be expressed in general as:
%
\begin{equation*}
\cclrd{s_1} \;\cpar\; (\cclrd{s_2} \,\cseq\, \cclrd{r}) \;\cgeq\;
  (\cclrd{s_1} \,\czip\, \cclrd{s_2}) \;\cpar\; ((\cclrd{s_1} \,\czip\, \cclrd{s_2}) \,\cseq\, \cclrd{r})
\end{equation*}
%
This property does not follow from the axioms of the flat coeffect algebra. To make
the flat coeffect system as general as possible, we do not \emph{in general} require it as
an additional axiom, although the above examples provide a reasonable basis for using
the specialized (\emph{let}) rule in the flat coeffect system.

@@

In the next section, we move from abstract treatment of the flat coeffect calculus to a more
concrete discussion. We explain its category-theoretical motivation, we use it to define
translational semantics (akin to Haskell's do notation) and we prove a soundness result that
well-typed programs in flat coeffect calcului for implicit parameters and dataflow do not get
stuck in the translated version.

>>>>

We enrich types and typing judgements with coeffect annotations $\cclrd{r}, \cclrd{s}, \cclrd{t}$;
typing judgements are written as $\coctx{\Gamma}{\cclrd{r}}\vdash e:\tau$. The expressions of
the calculus are those of the $\lambda$-calculus with \emph{let} binding. We also include a type
\ident{num} as an example of a concrete base type with numerical constants written as $n$:
%
\begin{equation*}
\begin{array}{rcl}
e &::=& x \sep n \sep \lambda x:\tau.e \sep e_1~e_2 \sep \kvd{let}~x = e_1~\kvd{in}~e_2\\
\tau &::=& \ident{num} \sep \tau_1 \xrightarrow{\cclrd{r}} \tau_2
\end{array}
\end{equation*}
%
Note that the lambda abstraction in the syntax is written in the Church-style and requires a type
annotation. This will be used in Section~\ref{sec:flat-unique} where we discuss how to find a
unique typing derivation for context-aware computations. Using Church-style lambda abstraction,
we can directly focus on the more interesting problem of finding unique \emph{coeffect annotations}
rather than solving the problem of type reconstruction.

@@

To guide understanding of the system, we also show non-syntax-directed (\emph{sub}) rule for
subcoeffecting. The rule states that an expression with context demands $\cclrd{r'}$ can be
treated as an expression with greater context demands $\cclrd{r}$. We return to subcoeffecting,
subtyping and additional constructs such as pairs in Section~\ref{sec:flat-exts}. When
discussing procedure for choosing unique typing in Section~\ref{sec:flat-unique}, we consider
only the syntax-directed part of the system.

@@

\paragraph{Properties.}
If a program written in a coeffect language with implicit parameters is well-typed in a type
system presented in Figure~\ref{fig:flat-resolve-impl} then this identifies the unique preferred
derivation for the program. We use this unique typing derivation to give the semantics of coeffect
language with implicit parameters in Chapter~\ref{ch:semantics} and we also implement this algorithm
as discussed in Chapter~\ref{ch:impl}.

@@

\paragraph{Properties.}
If a program written in a coeffect language for liveness or dataflow is well-typed according to
the type system presented in Figure~\ref{fig:flat-types} with the (\emph{abs}) rule replaced by
(\emph{idabs}), then the type system gives a unique derivation. As for implicit parameters, this
defines the semantics of coeffect program (Chapter~\ref{ch:semantics}) and it is used in the
implementation (Chapter~\ref{ch:impl}).

@@

\paragraph{Implicit parameters.} In dataflow, substituting $\kvd{prev}~x$ for a variable $y$ in
an expression $\lambda z.y$ changes the context demands attached to the type of the function.
This is the case not just for the preferred unique typing derivation, but for all possible typings
that can be obtained using the (\emph{abs}) rule. However, this is not the case for all
systems. Consider a substitution $\subst{\lambda x.y}{y}{\ident{?p}}$ that substitutes
an implicit parameter access $\ident{?p}$ for a free variable $y$ under a lambda:
%
\begin{equation*}
\begin{array}{rclcl}
 \coctx{y\!:\!\tau_1}{\emptyset} &\narrow{\vdash}& \lambda x.y : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{before}) \\
 \coctx{\emptyset}{ \{\ident{?p}\} } &\narrow{\vdash}& \lambda x.\ident{?p} : \tau_1 \xrightarrow{\emptyset} \tau_2 & & (\textnormal{after}) \\
\end{array}
\end{equation*}
%
The (\emph{after}) judgement shows one possible typing of the body -- one that does not change the
coeffects of the function type and attaches all additional coeffects (implicit parameters) to the
context. In case of implicit parameters (and, more generally, systems with set-like annotations)
this is always possible.

@@

It is perhaps somewhat unexpected that the annotation can be simplified in different ways for
different examples. However, for all our systems, the simplified annotation (right column in
Table~\ref{tab:flat-simplelet}) is more precise than the original (left column). Recall that
$\cclrd{s} = \cclrd{s_1}\,\czip\,\cclrd{s_2}$. The following inequalities hold:
%
\begin{equation*}
\begin{array}{rclll}
\cclrd{s_1} \cup (\cclrd{s_2} \cup \cclrd{r}) &\supseteq& (\cclrd{s_1} \cup \cclrd{s_2}) \cup \cclrd{r}
  && \textnormal{(implicit parameters)}\\
\cclrd{s_1} \sqcap (\cclrd{s_2} \sqcup \cclrd{r}) &\sqsupseteq&  (\cclrd{s_1} \sqcap \cclrd{s_2})
  && \textnormal{(liveness)} \\
\mathit{max}(\cclrd{s_1}, \cclrd{s_2} + \cclrd{r}) &\geq& \mathit{min}(\cclrd{s_1}, \cclrd{s_2}) + \cclrd{r}
  &\quad& \textnormal{(dataflow)} \\
\end{array}
\end{equation*}
%
In other words, the inequality states that using idempotence, we get a more precise typing.
Using the $\cgeq$ operator of flat coeffect algebra, this property can be expressed in general as:
%
\begin{equation*}
\cclrd{s_1} \;\cpar\; (\cclrd{s_2} \,\cseq\, \cclrd{r}) \;\cgeq\;
  (\cclrd{s_1} \,\czip\, \cclrd{s_2}) \;\cpar\; ((\cclrd{s_1} \,\czip\, \cclrd{s_2}) \,\cseq\, \cclrd{r})
\end{equation*}
%
This property does not follow from the axioms of the flat coeffect algebra. To make
the flat coeffect system as general as possible, we do not \emph{in general} require it as
an additional axiom, although the above examples provide a reasonable basis for using
the specialized (\emph{let}) rule in the flat coeffect system.

@@

In the next section, we move from abstract treatment of the flat coeffect calculus to a more
concrete discussion. We explain its category-theoretical motivation, we use it to define
translational semantics (akin to Haskell's ``do'' notation) and we prove a soundness result that
well-typed programs in flat coeffect calculi for implicit parameters and dataflow do not get
stuck in the translated version.
